---
title: "Analysis of the types of change present in Apache's sentry codebase over time"
author: "Fabián Santander"
date: "7/7/2021"
output: html_document
---

```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(ggplot2)
library(viridis)

# Load information and create RDA
if (file.exists("sentry_data.Rda")) {
  load("./sentry_data.Rda")
} else {
  
  sentry_data <- read_csv(quote = "'", file = "distill_results.csv")
  save(sentry_data, file="sentry_data.Rda")
  
}

# Generate cochanges_file_level dataframe if it doesn't exist and create RDA
if (file.exists("cochanges_file_level.Rda")) {
  load("./cochanges_file_level.Rda")
} else {
  
  cochanges_file_level <- sentry_data[ , c('filename', 'change', 'commit_1', 'commit_2')] %>%
  group_by(filename, commit_1, commit_2) %>%
  filter(n()==1) %>%
  group_by(commit_1, commit_2) %>%
  filter(n()>1)

  cochanges_file_level <- left_join(cochanges_file_level, cochanges_file_level, by=c('commit_1', 'commit_2'))
  
  cochanges_file_level <- cochanges_file_level[cochanges_file_level$filename.x != cochanges_file_level$filename.y,]
  
  # cochanges <- cochanges %>% filter(!duplicated(str_c(pmin(filename.x, filename.y), pmax(filename.x, filename.y) )))
  
  cochanges_file_level <- cochanges_file_level[, c('filename.x', 'change.x', 'filename.y', 'change.y', 'commit_2')]
  
  cochanges_file_level <- cochanges_file_level %>% rename(commit = commit_2)
  
  save(cochanges_file_level, file="cochanges_file_level.Rda")
  
}

# Generate cochanges_method_level dataframe if it doesn't exist and create RDA
if (file.exists("cochanges_method_level.Rda")) {
  load("./cochanges_method_level.Rda")
} else {
  
  cochanges_method_level <- sentry_data[ , c('level_name', 'change', 'change_level', 'commit_1', 'commit_2')] %>%
  group_by(level_name, commit_1, commit_2) %>%
  filter(n()==1) %>%
  group_by(commit_1, commit_2) %>%
  filter(n()>1)

  cochanges_method_level <- left_join(cochanges_method_level, cochanges_method_level, by=c('commit_1', 'commit_2'))
  
  cochanges_method_level <- cochanges_method_level[cochanges_method_level$level_name.x != cochanges_method_level$level_name.y,]
  
  # cochanges <- cochanges %>% filter(!duplicated(str_c(pmin(filename.x, filename.y), pmax(filename.x, filename.y) )))
  
  cochanges_method_level <- cochanges_method_level[, c('level_name.x', 'change.x', 'change_level.x', 'level_name.y', 'change.y', 'change_level.y', 'commit_2')]
  
  cochanges_method_level <- cochanges_method_level %>% rename(commit = commit_2)
  
  save(cochanges_method_level, file="cochanges_method_level.Rda")
  
}

# Global knitr options
knitr::opts_chunk$set(
  echo = FALSE,
  fig.width = 6,
  fig.height = 4,
  fig.align = 'center',
  knitr.table.format = "latex"
)

# Mask kable options to set global values
kbl <- function(data, caption, col.names=names(data)) {
  knitr::kable(data, booktabs = TRUE, digits = 3,caption = caption, col.names = col.names)
}

```

## Sentry data

This is a change analysis between files changed in commits of the project sentry by Apache. See: <https://github.com/apache/sentry>.

The data generated looks like:

```{r sentry data}
head(sentry_data, n = 20L) %>% 
  kbl("Sentry distilled commits") %>% 
  kable_styling() %>%
  scroll_box(height = "300px")
```

## Coupled changes

Over the period defined from commit `r head(sentry_data, 1)$commit_1` to `r tail(sentry_data, 1)$commit_2` whenever a pair of entities (files/methods/classes) were changed in the same commit, the pair and the commit on was recorded.

The cochanges dataset, at the file level, looks like:

```{r cochanges_file_level}
head(cochanges_file_level, n = 20L) %>% 
  kbl("Cochanges at file level") %>% 
  kable_styling() %>%
  scroll_box(height = "300px")
```

The cochanges dataset, at the method level, looks like:

```{r cochanges_method_level}
head(cochanges_method_level, n = 20L) %>% 
  kbl("Cochanges at method level") %>% 
  kable_styling() %>%
  scroll_box(height = "300px")
```

## Summary of data analyzed

Here's a brief summary of the data analyzed

### Types of changes

The frequencies of the type of changes observed overall are:

```{r type of changes, class.output="scroll-100"}

count(sentry_data, change, sort = TRUE) %>% 
  kbl("Frecuencies of type of changes") %>% 
  kable_styling() %>%
  scroll_box(height = "300px")

```

### Level of changes

The frequencies of the level of the changes made overall are:

```{r level of changes}

ggplot(sentry_data, aes(x=change_level, fill=change_level)) + 
  geom_bar(position="stack", stat="count") + 
  geom_text(stat='count', aes(label=..count..), vjust=-1) +
  scale_y_continuous(expand = c(0.2,0)) +
  scale_fill_viridis(discrete = T) +
  ggtitle("Levels of change", subtitle = "Apache's sentry codebase") + 
  ylab("Occurrences") +
  xlab("Level of change")

```

### Changes at the file level

Using the change information that we have available, we counted each change per file over the commits. The following info describes the occurrences of changes per file observed overall:

```{r summary of changes at file level}

changes_per_file <- count(sentry_data, filename, sort = TRUE)

changes_per_file %>%
  skimr::skim(n) %>%
  select(numeric.mean, numeric.sd, numeric.p0, numeric.p25, numeric.p50, numeric.p75, numeric.p100) %>%
  kbl(caption="Descriptive statistics of changes at file level", 
      col.names = c("mean","sd","p0","p25","p50","p75","p100")) %>%
  kable_styling()

```

These can be visualized using a box plot:

```{r boxplot of changes at file level}

boxplot(changes_per_file$n, main = "Boxplot of number of changes at file level", ylab = "Number of changes")

```

The outliers being:

``` {r outliers table at file level}

changes_per_file[changes_per_file$n > boxplot.stats(changes_per_file$n)$stats[5] ,] %>% 
  kbl(caption = "Table of outliers") %>% 
  kable_styling() %>%
  scroll_box(height = "300px")

```

### Changes at method/class/access level

Using the method level information available, we counted the changes recorded per method. The following info describes the occurrences of changes per method/class/access observed overall:

```{r summary of changes at method level}

changes_per_method <- count(sentry_data, level_name, sort = TRUE)

changes_per_method %>%
  skimr::skim(n) %>%
  select(numeric.mean, numeric.sd, numeric.p0, numeric.p25, numeric.p50, numeric.p75, numeric.p100) %>%
  kbl(caption="Descriptive statistics of changes at method level", 
      col.names = c("mean","sd","p0","p25","p50","p75","p100")) %>%
  kable_styling()

```

These can be visualized using a box plot:

```{r boxplot of changes at method level}

boxplot(changes_per_method$n, main = "Boxplot of number of changes at method level", ylab = "Number of changes")

```

The outliers being:

``` {r outliers table at method level}

changes_per_method[changes_per_method$n > boxplot.stats(changes_per_method$n)$stats[5] ,] %>% 
  kbl(caption = "Table of outliers") %>% 
  kable_styling() %>%
  scroll_box(height = "300px")

```

### Coupled changes

Here we used the info we got from the coupled changes analysis.

#### File level

Using the cochange information available, we counted the changes recorded per file. The following info describes the occurrences of changes per file observed overall:

```{r summary of cochanges at file level}

cochanges_per_file <- count(cochanges_file_level, filename.x, sort = TRUE)[, c('filename.x', 'n')]

cochanges_per_file[, c('filename.x', 'n')] %>%
  skimr::skim(n) %>%
  select(numeric.mean, numeric.sd, numeric.p0, numeric.p25, numeric.p50, numeric.p75, numeric.p100) %>%
  kbl(caption="Descriptive statistics of cochanges at file level", 
      col.names = c("mean","sd","p0","p25","p50","p75","p100")) %>%
  kable_styling()

```

These can be visualized using a box plot:

```{r boxplot of cochanges at file level}

boxplot(cochanges_per_file$n, main = "Boxplot of number of cochanges at file level", ylab = "Number of cochanges")

```

The outliers being:

``` {r outliers cochanges table at file level}

cochanges_per_file[cochanges_per_file$n > boxplot.stats(cochanges_per_file$n)$stats[5] ,] %>% 
  kbl(caption = "Table of outliers") %>% 
  kable_styling() %>%
  scroll_box(height = "300px")

```

#### Method level

Using the cochange information available, we counted the changes recorded per method. The following info describes the occurrences of changes per method observed overall:

```{r summary of cochanges at method level}

cochanges_per_method <- count(cochanges_method_level, level_name.x, sort = TRUE)[, c('level_name.x', 'n')]

cochanges_per_method[, c('level_name.x', 'n')] %>%
  skimr::skim(n) %>%
  select(numeric.mean, numeric.sd, numeric.p0, numeric.p25, numeric.p50, numeric.p75, numeric.p100) %>%
  kbl(caption="Descriptive statistics of cochanges at method level", 
      col.names = c("mean","sd","p0","p25","p50","p75","p100")) %>%
  kable_styling()

```

These can be visualized using a box plot:

```{r boxplot of cochanges at method level}

boxplot(cochanges_per_method$n, main = "Boxplot of number of cochanges at method level", ylab = "Number of cochanges")

```

The outliers being:

``` {r outliers cochanges table at method level}

cochanges_per_method[cochanges_per_method$n > boxplot.stats(cochanges_per_method$n)$stats[5] ,] %>% 
  kbl(caption = "Table of outliers") %>% 
  kable_styling() %>%
  scroll_box(height = "300px")

```



